---
title: "Wooldridge Ch2"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE}
load("C:/Users/Brendan/Desktop/Datasets/Wooldridge Econometrics Data/401k.RData")
```
## Ch 2 Answers

These are my answers to the end of chapter computer exercise C1 in Chapter 2 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C1
The variables in the dataset are:

1. prate - participation rate, percent
2. mrate - 401k plan match rate
3. totpart - total 401k participants
4. totelg - total eligible for 401k plan
5. age - age of 401k plan
6. totemp - total number of firm employees
7. sole - equal to 1 if 401k is firm's sole plan
8. ltotemp - log of totemp

###i)
```{r , echo=TRUE}
mean(data$prate)
mean(data$mrate)

```
###ii)

```{r , echo=FALSE}
plot(data$prate ~ data$mrate, xlab = "Match Rate", ylab = "Participation Rate")

```

```{r , echo = TRUE}
participation.model <- lm(prate ~ mrate, data = data)
summary(participation.model)
```
Sample size: 1534

R-squared: 0.075

###iii)
The intercept is 83.08 and the coefficient on mrate is 5.86. If the match rate is zero, the expected participation rate is 83.08%. The coefficient on mrate means that for each unit change in the match rate the expected participation rate changes by 5.86.

###iv)
```{r, echo=TRUE}
predict(participation.model, newdata = data.frame(mrate = 3.5))
```

This doesn't make sense because the participation rate cannot be greater than 100%. This is happening because we are using a model with no upper limit to predict an outcome with a limit at 100. The relationship does not appear to be linear so the model above is not appropriate.

###v)
The R-squared is very low at 0.075. The variation in the match rate does not explain much of the variation in the participation rate.
