---
title: "Wooldridge Ch2"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE}
load("C:/Users/Brendan/Desktop/Datasets/Wooldridge Econometrics Data/rdchem.RData")
```
## Ch 2 Answers

These are my answers to the end of chapter computer exercise C5 in Chapter 2 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C5

###i)

A constant elasticity model is a model in which the natural log is taken of both the dependent and independent variables before the regression is performed. This leads to a model where an approximate 1% change in the independent variable leads to an expected change of approximately $\beta_1$% in the dependent variable. In this model, the $\beta_1$ parameter is the elasticity.

\begin{equation}
ln(rd_i) = \beta_0 + \beta_1 ln(sales_i) + \epsilon_i
\end{equation}

###ii
```{r , echo=TRUE}
elasticity.model <- lm(log(rd) ~ log(sales), data = data)
summary(elasticity.model)
```
The estimated regression equation is,
\begin{equation}
\widehat{ln(rd_i)} = -4.1 + 1.08 ln(sales_i)
\end{equation}

The estimated elasticity of rd with respect to sales is 1.08. In other words, research and development is expected to increase by approximately 1.08% for every 1% increase in annual sales.
