---
title: "Wooldridge Ch3"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE}
load("C:/Users/Brendan/Desktop/Datasets/Wooldridge Econometrics Data/attend.RData")
```
## Ch 3 Answers

These are my answers to the end of chapter computer exercise C4 in Chapter 3 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C4
```{r, echo=FALSE}
desc
```
###i)

The minimum, maximum, and average of attendance rate, GPA prior to term, and ACT score to get the table below,

```{r, echo=FALSE}
atndrte <- c(min(data$atndrte), max(data$atndrte), mean(data$atndrte))
priGPA <- c(min(data$priGPA), max(data$priGPA), mean(data$priGPA))
ACT <- c(min(data$ACT), max(data$ACT), mean(data$ACT))
school <- matrix(c(atndrte,priGPA,ACT), ncol = 3)
colnames(school) <- c("atndrte", "priGPA","ACT")
rownames(school) <- c("min","max","average")
school <- as.table(school)
school
```

###ii)

The model, 
\begin{equation}
atndrte = \beta_0 + \beta_1*priGPA + \beta_2*ACT + u
\end{equation}

was estimated.
```{r, echo=FALSE}
attend.model <- lm(atndrte ~ priGPA + ACT, data=data)
summary(attend.model)
```
The estimated equation, sample size and $R^2$ are,

\begin{equation}
\widehat{atndrte} = 75.7 + 17.261*priGPA - 1.717*ACT \\
sample size = 680 \\
R^2 = 0.2906
\end{equation}

The intercept means that if a student had both prior GPA and ACT score equal to zero the expected attendance rate is about 76%. If it is the student's first term at the school then prior GPA could be zero but ACT score would not be zero so the intercept doesn't mean anything.

###iii)
The positive prior term GPA coefficient makes sense. If a student has a high GPA then they probably attend more classes. The negative ACT coefficient is a suprise though. A student with a high ACT score is likely hard working and studious and therefore more likely to attend classes. It's odd that a higher ACT score predicts lower class attendance.

###iv)
The predicted attendance rate of a student with a prior GPA of 3.65 and an ACT score of 20 is,
```{r, echo=FALSE}
predict(attend.model, newdata = data.frame(priGPA = 3.65, ACT = 20))
```
This result doesn't make sense because the maximum possible attendance rate is 100. A student can't attend more than 100% of the time.

There is a student with a prior GPA of 3.65 and ACT score of 20 and their attendance rate is 87.5. 

###v)
```{r, echo=FALSE}
A <- predict(attend.model, newdata = data.frame(priGPA = 3.1, ACT = 20))
B <- predict(attend.model, newdata = data.frame(priGPA = 2.1, ACT = 26))
difference <- A - B 

students <- matrix(round(c(A,B,difference),2), ncol = 3)
colnames(students) <- c("Student A", "Student B","Difference")
rownames(students) <- c("Attendance Rate")

students <- as.table(students)
students
```
The predicted difference between student A and student B is 27.6. The predicted attendance rate of student A is 94.9% and for student B is 67.3%.
