---
title: "Wooldridge Ch3"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE}
load("C:/Users/Brendan/Desktop/Datasets/Wooldridge Econometrics Data/bwght.RData")
```
## Ch 3 Answers

These are my answers to the end of chapter computer exercise C1 in Chapter 3 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C1

###i)
$\beta_2$ is most likely positive because we would expect higher income to allow access to better nutrition and care which would lead to a higher birthweight.

###ii
It is likely that cigs and faminc are negatively correlated because people tend to smoke less as income rises.
```{r, echo=TRUE}
plot(data$cigs ~ data$faminc)
cor(data$cigs,data$faminc)
```

There does appear to be a negative relationship between cigs and faminc in the above plot. The -0.173 correlation coefficient confirms this.

###iii)
```{r, echo=TRUE}
bwt.model.1 <- lm(bwght ~ cigs, data = data) 
bwt.model.2 <- lm(bwght ~ cigs + faminc, data = data)
summary(bwt.model.1)
summary(bwt.model.2)
```
The equation without faminc is,
\begin{equation}
\widehat bwght = 119.8 - 0.514 * cigs \\
samplesize = 1388 \\ 
R^2 = 0.023
\end{equation}

The equation with faminc included is,
\begin{equation}
\widehat bwght = 116.97 - 0.463 * cigs + 0.093 * faminc \\
samplesize = 1388 \\ 
R^2 = 0.0298
\end{equation}

The difference in the coefficient on cigs between the two models is about 0.05 ounces per cigarette. Since many of the women smoked 20 cigarettes per day this could result in the estimated birthweight differing by 1 ounce between the two models. This does not seem like a significant difference in practice and indicates that cigarette smoking has a similar negative effect on birthweight regardless of access to better prenatal care. 


