---
title: "Wooldridge Ch2"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE}
load("C:/Users/Brendan/Desktop/Datasets/Wooldridge Econometrics Data/wage2.RData")
```
## Ch 2 Answers

These are my answers to the end of chapter computer exercise C4 in Chapter 2 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C4


###i)
```{r , echo=TRUE}
mean(data$wage)
mean(data$IQ)
sd(data$IQ)
```

###ii)

```{r , echo=FALSE}
wage.model <- lm(wage ~ IQ, data = data)
summary(wage.model)
```

```{r , echo = TRUE}
wage.model$coefficients["IQ"]*15
```

The expected increase in monthly salary for an increase of 15 IQ points is about $125. With an R-squared of 0.096 the IQ score does not seem to explain much of the variation in monthly salary.

###iii)

```{r, echo=TRUE}
log.wage.model <- lm(log(wage) ~ IQ, data = data)
summary(log.wage.model)
```
```{r, echo=TRUE}
log.wage.model$coefficients["IQ"]*15
```
The log model predicts about a 0.9% increase in monthly salary for each one point increase in IQ. An increase of 15 IQ points would increase the expected monthly salary by about 13%.






