---
title: "Wooldridge Ch2 C3"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("C:/Users/Brendan/Desktop/Datasets/Wooldridge Econometrics Data/sleep75.RData")
```

## Ch 2 Answers

These are my answers to the end of chapter computer exercise C3 in Chapter 2 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C3

###i)
```{r, echo=TRUE}
sleep.model <- lm(sleep ~ totwrk, data = data)
summary(sleep.model)
```
The intercept is 3586.4 minutes of sleep per week and each additional minute of work results in a 0.15 minute drop in sleep. The intercept means that the expected amount of sleep for a person that does not work at all is 3586.4 minutes per week or about 8.5 hours a night on average. The R-squared is 0.1033 which means that the variation in the amount of sleep a person gets is not well explained by the variation in the amount of work they do. There seem to be other important factors that are not accounted for by the model.


###ii)
```{r, echo=TRUE}
#convert 2 hour increase to minutes
totwork.increase <- 2*60
#multiply the regression coefficient on total work by the increase in total work
sleep.decrease <- -0.15075*totwork.increase
sleep.decrease
```
If work over the course of the week increased by 2 hours (120 minutes) then the expected amount of sleep would decrease by about 18 minutes

This does not seem like a large effect, especially since the effect is weekly. This means that on average the amount of sleep lost for the additional 2 hours of work is only about 2.5 minutes a night.
