---
title: "Wooldridge Ch2"
author: "Brendan Peek"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1)
```
## Ch 2 Answers

These are my answers to the end of chapter computer exercise C8 in Chapter 2 of Introductory Econometrics 5e by Jeffrey Wooldridge.

###C8

###i)

```{r, echo=TRUE}
x <- runif(500,0,10)
mean(x)
sd(x)
```

The sample mean is 4.96 and the sample standard deviation is 2.83.

###ii
```{r, echo=TRUE}
u <- rnorm(500,0,6)
mean(u)
sd(u)
```

The sample average of u is not exactly zero because the sample is random but at this sample size it should almost always be close to zero. The sample standard deviation is 6.3 which is very close to what we would expect because the variance of the distribution from which the random sample was drawn is 36.

###iii)
```{r, echo=TRUE}
y <- 1 + 2*x + u
model <- lm(y~x)
summary(model)
```


###iv)
```{r, echo=TRUE}
sum(model$residuals)
sum(model$residuals*x)
```
The sum of the residuals and the sum of the product of the residuals and explanatory variables are both effectively zero. This result agrees with equation 2.60.

###v)
```{r, echo=TRUE}
sum(u)
sum(u*x)
```
These results definitely don't agree with equation 2.60.

###vi)
```{r, echo=TRUE}
set.seed(2)
x <- runif(500,0,10)
mean(x)
sd(x)

u <- rnorm(500,0,6)
mean(u)
sd(u)

y <- 1 + 2*x + u
model <- lm(y~x)
summary(model)
```
The intercept is quite a bit higher but the coefficient on x is about the same. They are different because of the different random samples used to estimate the model.
